\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}

% Raumbezogene Ingenieurwissenschaften = Geospatial Engineering

\begin{German}
    Diese Arbeit entstand im Rahmen meines Bachelorstudiums im Studiengang Raumbezogene Ingenieurwissenschaften. Um theoretische Inhalte mit praktischen Erfahrungen zu verknüpfen, suchte ich den Kontakt zu einem Industriepartner. Mein besonderer Dank gilt Axians für die Möglichkeit zur Zusammenarbeit. Namentlich danke ich Jean-Charles Schaegis für seine Unterstützung als praktischer Betreuer.

    Für die Vermittlung des Kontakts zum Institut für Bau- und Infrastrukturmanagement (IBI) danke ich Helena Laasch vom Institut für Geodäsie und Photogrammetrie (IGP). Ich danke Frau Prof. Catherine De Wolff dafür, dass sie die akademische Betreuung meiner Arbeit in ihrer Forschungsgruppe Circular Engineering for Architecture (CEA) ermöglichte. Die fachliche Betreuung erfolgte durch Dr. Kasimir Fort und Océane Durant-Maniclat, denen ich für ihre engagierte Unterstützung und konstruktiven Rückmeldungen besonders danke.

    Abschliessend danke ich Prof. Liangliang Nan von der TU Delft für seine Unterstützung bei der Integration der von ihm entwickelten Software.
\end{German}

\begin{English}
    This thesis was written as part of my bachelor's degree in Geospatial Engineering. To connect theoretical content with practical experience, I sought contact with an industry partner. My special thanks go to Axians for the opportunity to collaborate. I would like to thank Jean-Charles Schaegis in particular for his support as a practical supervisor.

    I would like to thank Helena Laasch from the Institute of Geodesy and Photogrammetry (IGP) for facilitating the contact with the Institute of Construction and Infrastructure Management (IBI). I am grateful to Prof. Catherine De Wolf for enabling the academic supervision of my work in her research group Circular Engineering for Architecture (CEA). The technical supervision was provided by Dr. Kasimir Forth and Océane Durand-Maniclas, to whom I extend my special thanks for their dedicated support and constructive feedback.

    Finally, I would like to thank Prof. Liangliang Nan from TU Delft for his support in integrating the software he developed.
\end{English}

\begin{flushright}
Jeffrey Leisi \\
Zurich, \the\year
\end{flushright}